---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fio

<!-- badges: start -->
[![R-CMD-check](https://github.com/albersonmiranda/fio/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/albersonmiranda/fio/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

A set of functions to help with input-output analysis, including international and brazilian matrices data, and a convenient addin for easy input.

The goal of fio is to provide a friedly interface to read in data from input-output matrices in Microsoft Excel to R. Additionally, it provides brazilian and international input-output matrices.

## Installation

You can install the development version of fio like so:

``` r
devtools::install_github("albersonmiranda/fio")
```

## Usage

Import complete brazilian 2020 input-output matrix:

```{r example-1}
iom = fio::iom_br_2020_51

# show 3 first row and columns from intermediate transactions matrix
iom[["Z"]][1:3, 1:3]
```

Import only intermediate transactions matrix from brazilian 2020 input-output matrix:

```{r example-2, eval=FALSE}
Z = fio::import_element(
  file = "excel_file.xlsx",
  sheet = "MIP",
  range = "D6:BB56",
  col_names = "D4:BB4",
  row_names = "D4:BB4"
)
```
