---
title: "Multi-Regional Input-Output Analysis"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multi-Regional Input-Output Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
p.caption {
  font-size: 0.6em;
  text-align: "center";
}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  out.width = "70%"
)
```

```{r setup}
library(fio)
```

# Introduction

Multi-Regional Input-Output (MRIO) analysis extends traditional input-output analysis to capture economic interdependencies between multiple regions or countries. The `miom` class in the `fio` package implements the methodology described in Miller & Blair (2009) for analyzing multi-regional economic systems.

This vignette demonstrates how to:

- Create and work with multi-regional input-output matrices
- Compute multi-regional multipliers
- Analyze inter-regional spillover effects  
- Extract country-specific input-output tables
- Measure regional interdependence

# Multi-Regional Input-Output Framework

In a multi-regional system with $R$ regions and $N$ sectors per region, the total dimensions of the system are $RN \times RN$. The intermediate transactions matrix $\mathbf{Z}$ is partitioned as:

$$\mathbf{Z} = \begin{bmatrix}
\mathbf{Z}^{11} & \mathbf{Z}^{12} & \cdots & \mathbf{Z}^{1R} \\
\mathbf{Z}^{21} & \mathbf{Z}^{22} & \cdots & \mathbf{Z}^{2R} \\
\vdots & \vdots & \ddots & \vdots \\
\mathbf{Z}^{R1} & \mathbf{Z}^{R2} & \cdots & \mathbf{Z}^{RR}
\end{bmatrix}$$

Where $\mathbf{Z}^{rs}$ represents intermediate flows from region $s$ to region $r$.

The multi-regional Leontief inverse is:

$$\mathbf{L} = (\mathbf{I} - \mathbf{A})^{-1}$$

Where $\mathbf{A}$ is the multi-regional technical coefficients matrix.

# Creating a Multi-Regional Input-Output Matrix

Let's create a 3-country, 3-sector example representing USA, China, and Germany with Agriculture, Manufacturing, and Services sectors.

```{r create_miom}
countries <- c("USA", "CHN", "DEU")
sectors <- c("Agriculture", "Manufacturing", "Services")

labels <- paste(rep(countries, each = 3), rep(sectors, 3), sep = "_")
labels

intermediate_transactions <- matrix(c(
  50, 20, 10,  5,  8,  2,  3,  4,  1,
  15, 80, 25,  8, 12,  5,  4,  6,  3,
  10, 30,120,  3,  7, 15,  2,  5, 10,
   4,  6,  2, 40, 15,  8,  2,  3,  1,
   7, 10,  5, 20, 70, 20,  5,  8,  4,
   5,  8, 12, 10, 25, 90,  3,  6,  8,
   2,  3,  1,  3,  4,  2, 30, 12,  5,
   4,  7,  3,  6,  9,  4, 15, 60, 15,
   3,  5,  8,  4,  6, 10,  8, 20, 80
), nrow = 9, ncol = 9, dimnames = list(labels, labels))

total_production <- matrix(c(200, 300, 450, 180, 280, 400, 150, 250, 350), 
                          nrow = 1, ncol = 9,
                          dimnames = list(NULL, labels))

intermediate_transactions[1:6, 1:6]
total_production
```

The intermediate transactions matrix shows the flows of goods and services between all country-sector combinations. Each row represents a purchasing sector, and each column represents a supplying sector. The total production vector contains the gross output for each country-sector combination.

```{r create_miom_object}
my_miom <- miom$new(
  id = "three_country_example",
  intermediate_transactions = intermediate_transactions,
  total_production = total_production,
  countries = countries,
  sectors = sectors
)

my_miom$n_countries
my_miom$n_sectors
dim(my_miom$intermediate_transactions)
```

The `miom` object inherits from the `iom` class, extending it with multi-regional functionality. It automatically stores information about the number of countries and sectors, and decomposes the transactions into domestic and international components.

# Multi-Regional Multiplier Analysis

The `miom` class computes several types of multipliers following Miller & Blair (2009):

- **Intra-regional multipliers**: Effects within the same region
- **Inter-regional multipliers**: Effects between different regions  
- **Spillover multipliers**: Total effects on other regions
- **Total multipliers**: Sum of intra-regional and spillover effects

```{r compute_multipliers}
my_miom$compute_multiregional_multipliers()

knitr::kable(my_miom$multiregional_multipliers, digits = 4)
```

## Interpreting the Multipliers

The multipliers table shows several key measures for each country-sector combination:

- **Intra-regional multiplier**: The total effect within the same country when that sector receives a $1 shock
- **Spillover multiplier**: The total effect on all other countries from the same shock  
- **Total multiplier**: The sum of intra-regional and spillover effects
- **Multiplier to [Country]**: The specific effect on each individual country

```{r interpret_multipliers}
usa_agr <- my_miom$multiregional_multipliers[
  my_miom$multiregional_multipliers$destination_label == "USA_Agriculture", ]

usa_agr[, c("intra_regional_multiplier", "spillover_multiplier", "total_multiplier", 
           "multiplier_to_USA", "multiplier_to_CHN", "multiplier_to_DEU")]
```

For example, when USA Agriculture receives a $1 increase in final demand, it generates $`r round(usa_agr$multiplier_to_USA, 4)` of total output in the USA, $`r round(usa_agr$multiplier_to_CHN, 4)` in China, and $`r round(usa_agr$multiplier_to_DEU, 4)` in Germany. The intra-regional effect is $`r round(usa_agr$intra_regional_multiplier, 4)`, while the spillover to other countries totals $`r round(usa_agr$spillover_multiplier, 4)`.

# Regional Interdependence Analysis

Regional interdependence measures help understand the economic integration between countries. The `get_regional_interdependence()` method computes several key indicators:

- **Self-reliance**: Average intra-regional multiplier (higher values indicate more self-sufficient economies)
- **Spillover out**: Average impact on other regions when this country is shocked
- **Spillover in**: Average sensitivity to shocks from other regions  
- **Interdependence index**: Ratio of spillover to self-reliance

```{r interdependence}
interdependence <- my_miom$get_regional_interdependence()
knitr::kable(interdependence, digits = 4)
```

The results show that `r interdependence$country[which.max(interdependence$self_reliance)]` has the highest self-reliance (`r round(max(interdependence$self_reliance), 4)`), indicating a more self-sufficient economy. Meanwhile, `r interdependence$country[which.max(interdependence$interdependence_index)]` has the highest interdependence index (`r round(max(interdependence$interdependence_index), 4)`), suggesting greater integration with other economies.

# Bilateral Trade Analysis

The `miom` class allows extraction of bilateral trade flows between any two countries using the `get_bilateral_trade()` method:

```{r bilateral_trade}
usa_to_chn_trade <- my_miom$get_bilateral_trade("USA", "CHN")
knitr::kable(usa_to_chn_trade, digits = 2, 
             caption = "Trade flows from USA to China (by sector)")

chn_to_deu_trade <- my_miom$get_bilateral_trade("CHN", "DEU")
knitr::kable(chn_to_deu_trade, digits = 2,
             caption = "Trade flows from China to Germany (by sector)")
```

These matrices show the intermediate goods flows from the origin country (columns) to the destination country (rows) by sector. The values represent the monetary flows of intermediate inputs used in production.

# Net Spillover Effects

Net spillover effects reveal asymmetric economic relationships between countries. Positive values indicate that the row country benefits more from economic activity in the column country than vice versa.

```{r net_spillover}
net_spillover <- my_miom$get_net_spillover_matrix()
knitr::kable(net_spillover, digits = 4, caption = "Net Spillover Effects Matrix")
```

The interpretation is straightforward:
- Positive values: row country benefits more from column country
- Negative values: column country benefits more from row country  
- Values close to zero: symmetric spillover relationship

```{r max_spillover}
max_abs_spillover <- which(abs(net_spillover) == max(abs(net_spillover)), arr.ind = TRUE)
max_row <- rownames(net_spillover)[max_abs_spillover[1]]
max_col <- colnames(net_spillover)[max_abs_spillover[2]]
max_value <- net_spillover[max_abs_spillover]
```

The strongest asymmetric relationship is from `r max_col` to `r max_row` with a net spillover of `r round(max_value, 4)`.

# Extracting Country-Specific Input-Output Tables

The `extract_country()` method allows you to extract domestic input-output tables for individual countries from the multi-regional system:

```{r extract_country}
usa_iom <- my_miom$extract_country("USA")

usa_iom$id
dim(usa_iom$intermediate_transactions)
```

The extracted `iom` object contains only the domestic transactions for the specified country:

```{r usa_tables}
knitr::kable(usa_iom$intermediate_transactions, digits = 2,
             caption = "USA Domestic Intermediate Transactions")

knitr::kable(usa_iom$total_production, digits = 2,
             caption = "USA Total Production")
```

You can then perform standard input-output analysis on the domestic economy:

```{r usa_multipliers}
usa_iom$compute_tech_coeff()$compute_leontief_inverse()$compute_multiplier_output()

knitr::kable(usa_iom$multiplier_output, digits = 4,
             caption = "USA Domestic Output Multipliers")
```

These domestic multipliers capture only the within-country effects, excluding international spillovers that would be present in the full multi-regional analysis.

# Advanced Analysis: Spillover Matrix

The spillover matrix provides a comprehensive view of how shocks in each country-sector affect all other country-sectors in the system:

```{r spillover_matrix}
spillover_matrix <- my_miom$get_spillover_matrix()
dim(spillover_matrix)
```

This `r dim(spillover_matrix)[1]` × `r dim(spillover_matrix)[2]` matrix contains the complete set of multiplier effects. For example, we can examine how a shock to USA Agriculture affects other countries:

```{r usa_agr_effects}
usa_agr_effects <- spillover_matrix[, "USA_Agriculture"]
other_country_effects <- usa_agr_effects[!grepl("^USA_", names(usa_agr_effects))]
head(other_country_effects, 6)
```

These values show the output response in each foreign country-sector to a unit shock in USA Agriculture.

# Key Sectors Analysis

Key sectors analysis identifies sectors with strong backward and forward linkages in the multi-regional system:

```{r key_sectors}
my_miom$compute_key_sectors()

key_sectors_table <- my_miom$key_sectors[, c("sector", "country", "sector_name", 
                                            "power_dispersion", "sensitivity_dispersion")]
knitr::kable(key_sectors_table, digits = 4, caption = "Key Sectors Analysis")
```

Key sectors are typically defined as those with above-average values in both power dispersion (backward linkages) and sensitivity dispersion (forward linkages):

```{r identify_key_sectors}
mean_power <- mean(my_miom$key_sectors$power_dispersion)
mean_sensitivity <- mean(my_miom$key_sectors$sensitivity_dispersion)

key_sectors_identified <- my_miom$key_sectors[
  my_miom$key_sectors$power_dispersion > mean_power & 
  my_miom$key_sectors$sensitivity_dispersion > mean_sensitivity, 
  c("sector", "country", "sector_name")
]

if (nrow(key_sectors_identified) > 0) {
  knitr::kable(key_sectors_identified, caption = "Identified Key Sectors")
} else {
  NULL
}
```

`r if (nrow(key_sectors_identified) > 0) { paste("The analysis identifies", nrow(key_sectors_identified), "key sectors with above-average linkages in both dimensions.") } else { "No sectors meet the criteria for key sectors with the current thresholds, suggesting a relatively balanced linkage structure across the multi-regional system." }`

# Summary and Conclusions

This vignette demonstrated the key functionalities of multi-regional input-output analysis using the `miom` class:

1. **Multi-regional multipliers**: Quantify how economic shocks propagate within and across regions
2. **Interdependence measures**: Assess the degree of economic integration between countries
3. **Bilateral trade analysis**: Extract specific trade relationships between country pairs
4. **Spillover effects**: Measure asymmetric economic relationships
5. **Country extraction**: Analyze domestic economies within the multi-regional framework
6. **Key sectors identification**: Find sectors with strong backward and forward linkages

## Key Insights from the Example

```{r summary_insights}
avg_intra <- mean(my_miom$multiregional_multipliers$intra_regional_multiplier)
avg_spillover <- mean(my_miom$multiregional_multipliers$spillover_multiplier)
most_self_reliant <- interdependence$country[which.max(interdependence$self_reliance)]
max_self_reliance <- max(interdependence$self_reliance)
most_interdependent <- interdependence$country[which.max(interdependence$interdependence_index)]
max_interdependence <- max(interdependence$interdependence_index)
```

The analysis reveals several important characteristics of our three-country system:

- The average intra-regional multiplier is `r round(avg_intra, 4)`, indicating the typical within-country effect of economic shocks
- The average spillover multiplier is `r round(avg_spillover, 4)`, showing the extent of cross-border economic integration
- `r most_self_reliant` exhibits the highest self-reliance (`r round(max_self_reliance, 4)`), suggesting stronger domestic production chains
- `r most_interdependent` shows the highest interdependence index (`r round(max_interdependence, 4)`), indicating greater sensitivity to international economic conditions

These measures provide valuable insights for understanding regional economic integration and the potential impacts of economic policies or external shocks across the multi-regional system.

## References

Miller, R. E., & Blair, P. D. (2009). *Input-output analysis: foundations and extensions*. Cambridge University Press.

The methodology implemented in the `miom` class follows the multi-regional input-output framework described in Chapter 3 of Miller & Blair (2009), with extensions for spillover analysis and regional interdependence measures.