---
title: "benchmarking"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This vignette presents a benchmarking analysis comparing the performance of functions from the `fio` package with equivalent base R functions. The `fio` package provides a set of functions for input-output analysis, a method used in economics to analyze the interdependencies between different sectors of an economy.

In this document, we will focus on two key functions: the technical coefficients matrix calculation and the Leontief inverse matrix calculation. These functions are fundamental to input-output analysis, and their performance can significantly impact the speed of larger analyses.

Our benchmarking tests, which involve running these functions repeatedely in simulated datasets, show that the `fio` package functions are faster than the equivalent base R functions. This improved performance can make a substantial difference in larger analyses, making the `fio` package a valuable tool for input-output analysis in R.

The tests were run on a simulated $1000 \times 1000$ matrix, and each test was repeated 100 times to account for variability. Please note that the results of this benchmarking analysis are dependent on the specific test datasets used and the hardware on which the algorithms are run. Therefore, the results should be interpreted in the context of these specific conditions.

## Technical coefficients matrix

The technical coefficients matrix calculation, a key step in input-output analysis, was tested using the `tec_coeff()` function from the `{fio}` package, equivalent functions from the `{leontief}` package, and a base R implementation. The results show that both `{fio}` and `{leontief}` functions execute almost instantaneously, with `{leontief}` slightly faster. In contrast, the base R implementation is about 100 times slower.

```{r benchmark_a}
# set seed
set.seed(100)

# data
matrix_dim <- 2000
intermediate_transactions <- matrix(
  as.double(sample(1:1000, matrix_dim^2, replace = TRUE)),
  nrow = matrix_dim,
  ncol = matrix_dim
)
total_production <- matrix(
  as.double(sample(4000000:6000000, matrix_dim, replace = TRUE)),
  nrow = 1,
  ncol = matrix_dim
)

# Base R function
tec_coeff_r <- function(intermediate_transactions, total_production) {
  tec_coeff_matrix <- intermediate_transactions %*% diag(1 / as.vector(total_production))
  return(tec_coeff_matrix)
}

# {fio} setup
iom_fio <- fio::iom$new("iom", intermediate_transactions, total_production)

# benchmark
benchmark_a <- microbenchmark::microbenchmark(
  `Base R` = tec_coeff_r(intermediate_transactions, total_production),
  leontief = leontief::input_requirement(intermediate_transactions, total_production),
  fio = fio:::tec_coeff(intermediate_transactions, total_production),
  times = 100,
  check = "equivalent"
)
print(benchmark_a)

# plot
ggplot2::autoplot(benchmark_a)
```

## Leontief inverse matrix

For the Leontief inverse matrix calculation, we will use the technical coefficients matrix calculated in the previous section. Here, `{fio}` is the fastest algorithm, followed by `{leontief}` and a surprisingly efficient base R implementation. 

```{r benchmark_b}
# data
iom_fio <- iom_fio$tec_coeff()
technical_coefficients_matrix <- iom_fio$technical_coefficients_matrix

# R function
leontief_inverse_r <- function(technical_coefficients_matrix) {
  dim <- nrow(technical_coefficients_matrix)
  leontief_inverse_matrix <- solve(diag(dim) - technical_coefficients_matrix)
  return(leontief_inverse_matrix)
}

# benchmark
benchmark_b <- microbenchmark::microbenchmark(
  `Base R` = leontief_inverse_r(technical_coefficients_matrix),
  leontief = leontief::leontief_inverse(technical_coefficients_matrix),
  fio = fio:::leontief_inverse(technical_coefficients_matrix),
  times = 100,
  check = "equivalent"
)
print(benchmark_b)

# plot
ggplot2::autoplot(benchmark_b)
```
